TRIK Studio: Technical Introduction

### 2017-02-10 2017-trikStudioEnglish/trikStudioEnglish.tex
Polishing introduction

\author{
	\IEEEauthorblockN{Dmitry Mordvinov}
	\IEEEauthorblockA{
		Department of Software Engineering \\
		St.-Petersburg State University \\
		St.Petersburg, Russia \\
		Email: mordvinov.dmitry@gmail.com
	}
\and
	\IEEEauthorblockN{Yurii Litvinov}
	\IEEEauthorblockA{
		Department of Software Engineering \\
		St.-Petersburg State University \\
		St.Petersburg, Russia \\
		Email: y.litvinov@spbu.ru 
	}
\and
	\IEEEauthorblockN{Timofey Bryksin}
	\IEEEauthorblockA{
		Department of Software Engineering \\
		St.-Petersburg State University \\
		St.Petersburg, Russia \\
		Email: t.bryksin@spbu.ru 
	}
}

@@

Almost all aforementioned visual languages are based on control-flow computational model. Such a model is easier to understand, so it suits the purpose of education well. Nevertheless this model is not the best for programming robots since they are highly reactive by their nature: a program controlling a robot is basically a transformation of sensor signals into motor impulses. Reactive models are best expressed in data-flow languages~\cite{johnston2004advances}, where a program consists of a number of <<black boxes>>, connected with data channels. Each such <<box>> (we will call it \textit{block}) has a fixed number of inputs and outputs, and its job is to transform input data into output data (we will call \textit{tokens} data transferring through channels). For instance, each robot's sensor is represented by a single block, that simply sends tokens with sensor data to its output. Many researchers  note usability of visual data-flow languages comparing to control-flow ones~\cite{johnston2004advances}, in particular because of clear visualization of data flows.

The idea of using data-flow programming languages in robotics is not a novel one, they are implemented in almost every toolset for programming industrial and laboratory automation system. Among them are LabVIEW by National Instruments~\cite{kodosky1991visual}, Simulink~\cite{dabney2004mastering}, and Microsoft Robotics Developer Studio~\cite{jackson2007microsoft}. All of them solve the automation task very well and have very powerful tools inside (for example, Microsoft Robotics Developer Studio is run on MySpace social network servers~\cite{scherotter2009ccr}), but are very complex and hard to learn, so even if they are used in education, only in universities (\cite{stefanovic2011labview,yi2005labview}). School experiments with LabVIEW were popular in late 1990s (\cite{cyr1997low,portsmore1999robolab}), which resulted in adaptations of this environment (like Robolab), where data-flow model was replaced with control-flow model. So, a gap between simplified languages suitable for education and more elegant, but complex industrial languages exists. Some researchers are trying to adapt the data-flow model for educational robotics, for example a research group from New Zealand introduced the RuRu~\cite{diprose2011ruru} language. Nevertheless to our best knowledge there are still no production-ready data-flow based programming environments for educational purpose today.

Current paper describes TRIK Studio programming environment, an attempt to solve problems mentioned above. The success in solving them is indirectly acknowledged by its applications: currently TRIK Studio is widely used in Russian education organizations (almost a hundred of schools and robotic clubs), several cases of TRIK Studio usage in organizations in Great Britain and France are known, there are TRIK Studio users on every inhabited continent. This paper presents technical overview of the programming environment, educational and methodical issues are not discussed in detail.

>>>>

\author{
	\IEEEauthorblockN{Dmitry Mordvinov}
	\IEEEauthorblockA{
		Department of Software Engineering \\
		St.-Petersburg State University \\
		St.Petersburg, Russia \\
		Email: dmitry.mordvinov@se.math.spbu.ru
	}
\and
	\IEEEauthorblockN{Yurii Litvinov}
	\IEEEauthorblockA{
		Department of Software Engineering \\
		St.-Petersburg State University \\
		St.Petersburg, Russia \\
		Email: y.litvinov@spbu.ru 
	}
\and
	\IEEEauthorblockN{Timofey Bryksin}
	\IEEEauthorblockA{
		Department of Software Engineering \\
		St.-Petersburg State University \\
		St.Petersburg, Russia \\
		Email: t.bryksin@spbu.ru 
	}
}

@@

Almost all aforementioned visual languages are based on control-flow computational model. Such a model is easier to understand, so it suits the purpose of education well. Nevertheless this model is not the best for programming robots since they are highly reactive by their nature: a program controlling a robot is basically a transformation of sensor signals into motor impulses. Reactive models are best expressed in data-flow languages~\cite{johnston2004advances}, where a program consists of a number of <<black boxes>>, connected with data channels. Each such <<box>> (we will call it \emph{block}) has a fixed number of inputs and outputs, and its job is to transform input data into output data (we will call \emph{tokens} data transferring through channels). For instance, each robot's sensor is represented by a single block, that simply sends tokens with sensor data to its output. Many researchers  note usability of visual data-flow languages comparing to control-flow ones~\cite{johnston2004advances}, in particular because of clear visualization of data flows.

The idea of using data-flow programming languages in robotics is not a novel one, it was adopted in almost every toolset for programming industrial and laboratory automation systems, such as LabVIEW by National Instruments~\cite{kodosky1991visual}, Simulink~\cite{dabney2004mastering} and Microsoft Robotics Developer Studio~\cite{jackson2007microsoft}. They all have powerful backends (for example, Microsoft Robotics Developer Studio is being used on MySpace social network servers~\cite{scherotter2009ccr}) but are very complex and hard to learn. Thus they are used only in higher education~(\cite{stefanovic2011labview,yi2005labview}). School experiments with LabVIEW were popular in late 1990s~(\cite{cyr1997low,portsmore1999robolab}), which resulted in adaptations (like Robolab) replacing data-flow model  with control-flow model. So, there is a gap between simplified languages suitable for education and more elegant but complex industrial languages. There are attempts to adapt a data-flow model for educational robotics, for example a research group from New Zealand proposed the RuRu~\cite{diprose2011ruru} language. Nevertheless, to the best of our knowledge, there is still no production-ready data-flow-based programming environment suitable for secondary education today.

Current paper describes TRIK Studio programming environment, which attempts to solve the aforementioned problems. TRIK Studio is widely used in Russian education organizations (almost a hundred of schools and robotic clubs), several applications in Great Britain, France and Singapore are known. This paper presents technical overview of the programming environment; educational and methodical issues are not discussed in detail.
